---
title: "Exercises"
output: html_document
date: "2024-11-06"
---

Exercise 1: Summary of the Paper and Main Findings
Provide a brief summary of the paper you are replicating. Describe the main findings, especially
those related to Study 1.

Disclaimer: For this exercise I made use of the following AI Generation tools:
- Deepl.com for translation purposes
- The integrated Copilot tool in R Studio for generating code and text
- The free versions of OpenAI's GPT-4o Mini, Anthropics Claude 3.5 Sonnet and Mistral's Codestral AI model for generating code and assisting me in solving the tasks.

The authors posit that according to the Homonationalist thesis, support for LGBT+ issues among anti-immigration nativists and right-wing populists is not entirely genuine (organic liberalism thesis) and is instead used strategically (instrumental liberalism thesis) as a tool to justify further exclusion of immigrants, especially Muslims, who are perceived to be more homophobic, than the average western citizen. The rational behind this could be summed up as: "The enemy of my enemy is my friend" or: "I hate them (Muslims) more than I hate the others (LGBT+ people)". 
In order to test this hypothesis, the authors conducted two survey experiments in the UK and Spain. The dependent variable is "Support for LGBT+ inclusive education in school" and was measured by means of survey response on an order randomized scale from 0 to 11, but then for the sake of simplicity were turned into a dichotomous variable for support (0 or 1). 
In the UK, for study 1, the experimental research design randomly assigned 1200 participants into a control group and a treatment group. Those in the control group were exposed to text and pictures of nondenominational people with conventional British names protesting against LGBT+ inclusive education in schools. The treatment group was exposed to the same pictures and text, but the names and pictures were changed to appear Muslim. 
The authors expected that individuals who harbor negative predispositions toward immigration in the treatment group, will display more socially liberal preferences toward LGBT+ education than individuals with comparable anti-immigrant sentiment in the control group.
Indeed, the results showed that (with an average treatment effect of 9.7% points or a 22% increase against the baseline) respondents with an anti-immigrant disposition in the treatment group showed increased support for LGBT+ inclusive education in schools than comparable individuals in the control group. 
On the other hand, among those with positively predisposed toward immigration, treatment exhibits no significant effect below the 5% significance level.
But overall, it can be said, that support for LGBT+ inclusive education is positively correlated with liberal attitudes towards immigrants.
In Study 2, conducted in Spain, chosen because of generally more positive attitude levels towards immigration and LGBT+ issues than in the UK, and a right-wing populist party (VOX), that unlike its counter-part in the UK (UKIP), is not following a homonationalist strategy and is more openly outspoken against LGBT+ issues. They found that regardless of immigration sentiments, treatment resulted in an 10-11% increase in support for LGBT+ inclusive education in schools. However, the effect was stronger among those with anti-immigrant sentiments (21% vs. 14%), when compared against the baseline.
Furthermore, the authors also looked at shifts in citizens pride in "Western Culture". Only among individuals with nativists views could they reasonably demonstrate increased levels of "pride in Western Culture" in the treatment group.

With these findings in mind, the authors argue that the homonationalist thesis is correct and that the increased levels of support for LGBT+ issues among the wider general public in western countries is not entirely genuine, but also in part the result of a strategic move by right-wing populists to deradicalize their image by defining themselves by "who they are not" rather than "who they are". They further speculate that as soon as the political context changes and LGBT+ rights are at stake (but absent of the perceived threat of "immigrants"), the support for LGBT+ issues among right-wing populists will decrease. 

Exercise 2: Data Preparation and Exploration
Use the data file study1_data.csv to begin the replication process. Identify and describe the
experimental variables (i.e., treatment and immigration attitudes) and provide visualizations of their distribution.
Then, select up to four covariates (e.g., gender, age, etc.) and plot their distribution too.
If necessary, clean or transform variables. Document any changes.


From looking at the csv table for Study1 as well as looking at the study1.R script, one can deduce which variables are likely of interest for the replication. 
As a first step, I need to find the dependent variable (DV) support for LGBT+ inclusive education, which according to the text had a 0-11 range with a randomized scale order that was transformed to 0 and 1 for easier interpretation.
Secondly, I need to find the experimental variables for treatment and immigration attitudes.
The treatment variable concerns the exposure to Muslim names and pictures and is most likely operationalized as a binary variable (0,1).
The variable for immigration attitudes is likely to be operationalized as an ordinal scale (0-10) or as a binary variable.

A look at the Csv table revealed the following variables as of interest for the replication:
- Treatment variable: outcome_treat (categorical or nominal, with NAs), treatment (binary),
- Immigration attitudes: imm_1, imm_2, (both likely ordinal or numeric) imm1mean (continous, same value), immbelow (binary), imm3 (ordinal) 
- DV: support (binary), support2 (categorical)

By also looking at the study1.R script in the code chunk below, I can see that the following variables are of interest for the replication, while the rest can likely be ignored:

```{r}
df <-df%>% 
  mutate(treat= as.factor(treatment), --> convert to factor (categorical data)
         treatnum= as.numeric(treatment), --> convert to numeric
         #[...]
         immbelow= as.factor(immbelow),
         imm3= as.factor(imm3),
         
##Figure 3## 

model1<- glm (support ~ treat*imm_1, data=df, family="binomial")
summ(model1, robust=TRUE)
df$predictINT1<-predict(model1, df, type="response")

treat <- subset(df, treatnum==1)
control <- subset(df, treatnum==0)
proimm <- subset(df, immbelow==0)
noproimm <- subset(df, immbelow==1)

##FIGURE 4##

modelsub1<- lm (support ~ treat, data=proimm)
summ(modelsub1, robust=TRUE)
proimm$predictedb<-predict(modelsub1, proimm)

modelsub2<- lm (support ~ treat, data=noproimm)
summ(modelsub2, robust=TRUE)
noproimm$predictedb<-predict(modelsub2, noproimm)

treatsub1 <- subset(proimm, treatnum==1)
controlsub1 <- subset(proimm, treatnum==0)
treatsub2 <- subset(noproimm, treatnum==1)
controlsub2 <- subset(noproimm, treatnum==0)
```

- Treatment variable: treatment: is transformed twofold -> treat (converted to cat. variable for model transformation), treatnum (converted to numeric binary variable for graph)
- Immigration attitudes: immbelow (cat., converted to binary variable for graph), imm_1 (ordinal, 0-10 with 10 standing for maximal positive attitude towards immigrants)
- DV: support (binary)

From this I conclude that the following variables are of interest for the replication:
- Treatment Variable Operationalization: treatment (0,1)

- Immigrant Attitudes Operationalization: imm_1 (ordinal scale 0-10) ((or immbelow, for binary distribution Pro-/AntiImmigration))

- DV operationalization: support (0,1)

Below I will provide visualizations of the distribution of these variables. Let's start with the treatment variable:
```{r}
#Loading Data Set
study1 <- read.csv("/Users/nicolaswaser/New-project-GitHub-first/R/Take-Home-Ex.-1/Input Data/study1_data.csv",)

# Get idea of variable distribution in the form of a table
table(study1$treatment)
```

```{r}
#Visualizing the distribution of the treatment variable
ggplot(study1, aes(x=treatment)) + geom_bar() + labs(title="Distribution of Treatment Variable", x="Treatment", y="Frequency")
```
Unsurprisingly, the treatment variable is distributed evenly across the two groups, as it should be in a randomized experiment. 
Next, I will visualize the distribution of the Immigration attitude variable imm_1:
```{r}
table(study1$imm_1)
```

```{r}
#Visualizing the distribution of imm_1
ggplot(study1, aes(x=imm_1)) + geom_bar() + labs(title="Distribution of attitudes towards immigrants", x="Attitudes towards immigrants (from negative to positive)", y="Frequency")
```
The distribution of attitudes towards immigrants is skewed towards the positive side, with most respondents having a positive attitude. Note: 0 values equal to NAs, which is why the distribution starts at 1.
For later purposes, I will also visualize the distribution of the binary variable immbelow:
```{r}
#Visualizing the distribution of immbelow
ggplot(study1, aes(x=immbelow)) + geom_bar() + labs(title="Distribution of Pro- vs. Anti-immigration attitudes", x="Pro- (0) vs. Anti- (1) immigration", y="Frequency")
```
There are slightly more Pro-immigration than Anti-immigration sentiments in the sample. This is also in line with the graph from above. 
Lastly, I will visualize the distribution of the DV support for LGBT+ inclusive education:
```{r}
#Visualizing the distribution of support for LGBT+ inclusive education
ggplot(study1, aes(x=support)) + geom_bar() + labs(title="Distribution of support for LGBT+ inclusive education", x="Supports LGBT+ inclusive education, No (0), Yes (1)", y="Frequency")
```
```{r}
# Distribution of support 2
ggplot(study1, aes(x=support2)) + geom_bar() + labs(title="Distribution of support for LGBT+ inclusive education", x="Stance on LGBT+ inclusive education", y="Frequency")
```
The distribution of support for LGBT+ inclusive education is skewed towards the positive side, with most respondents supporting LGBT+ inclusive education. Just to be sure that 0 stands for opposition and 1 for support, I also visualized the distribution of support2.

Selecting covariates:
I pick the following covariates for the analysis:
- age: Age is often associated with a change in political attitudes over time -> age (numeric) or agecat (categorical) -> I pick age so that I have a continuous variable in the analysis
- gender: Gender is associated with different attitudes towards LGBT+ rights and immigration and could impact that relationship as well -> I pick gender (categorical, Man, Woman), as it is a binary variable and make things less complicated
- sexual identity/orientation: I expect that identifying as LGBT+ will have a big influence on the relationship of attitudes towards LGBT+ inclusive education and exposure to the treatment-> I pick the variable queer (categorical, Heterosexual, Not heterosexual) as it is a binary variable and makes things less complicated.
- race: not being white (in the case of the UK) might result in different attitudes towards LGBT+ inclusive education -> I pick the variable nonwhite.

Age:
```{r}
#Visualizing the distribution of age
ggplot(study1, aes(x=age)) + geom_histogram(binwidth = 1) + scale_x_continuous() + labs(title="Distribution of age", x="Age", y="Frequency"))
```
```{r}
#Visualizing the distribution of agecat
ggplot(study1, aes(x=agecat)) + geom_bar() + labs(title="Distribution of age", x="Age Cohort", y="Frequency")
```
The distribution of age is skewed towards older respondents, with most respondents being in their 30s and above. Note: 

Gender:
```{r}
#Visualizing the distribution of Gender (Man, Woman)
ggplot(study1, aes(x=gender)) + geom_bar() + 
  labs(title="Distribution of gender", x="Gender", y="Frequency")
```
There are slightly more Men than Women in the sample, but overall they are distributed evenly.

Sexual Identity:
```{r}
#Visualizing the distribution of queer
ggplot(study1, aes(x=queer)) + geom_bar() + 
  labs(title="Distribution of sexual identity", x="Sexual Identity", y="Frequency")
```
The large majority of respondents identify as heterosexual, with only a small minority identifying as not heterosexual.

Data Cleaning/transformation if necessary:



Exercise 3: Covariate Balance
Check for balance across covariates and report the results in a table and a plot.
Explain your findings. Why would you expect randomization to lead to balance across covariates?

```{r}
#Checking for balance across covariates
# Load necessary libraries
library(dplyr)
library(ggplot2)
install.packages("tableone")
library(tableone)

# Assuming the dataset is named `study1_data`
# and has columns `treatment` (1 = treated, 0 = control),
# `gender`, `age`, and `queer`

# Step 1: Create a Balance Table
# Convert treatment to a factor for easier comparison
study1$treatment <- as.factor(study1$treatment)

# Define covariates and treatment
covariates <- c("gender", "age", "queer")
treatment <- "treatment"

# Generate a summary table for covariate balance
balance_table <- CreateTableOne(vars = covariates, strata = treatment, data = study1, test = FALSE)
print(balance_table)

# Step 2: Calculate Standardized Mean Differences (SMD)
# Extract SMDs from balance_table
smd_values <- ExtractSmd(balance_table)
print(smd_values)

### length names are not the same
### error note!

length(names(smd_values))
length(unlist(smd_values))

# Step 3: Plotting Balance with Standardized Mean Differences
# Converting SMDs to a data frame for plotting
smd_df <- data.frame(
  covariate = names(smd_values),
  smd = unlist(smd_values)
)

# Plotting the SMDs
ggplot(smd_df, aes(x = covariate, y = smd)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "red") +
  labs(title = "Covariate Balance Check using Standardized Mean Differences",
       y = "Standardized Mean Difference (SMD)", x = "Covariates") +
  theme_minimal()
```


Exercise 4 (additional): Estimate Treatment Effect
This exercise is not mandatory, but it serves only to opt for the maximum grade (6).
Estimate the average effect of the treatment on the outcome variable support conditional on
the pre-treatment immigration attitudes imm_1. For this, use an interaction term in an OLS
regression model. Compare your results to those in the original paper.
Then, repeat this analysis with three iteratively smaller random samples of the treatment 
(n = 200, n = 100, n = 10) and control groups (n = 200, n = 100, n = 10); total N = 400, 200, and 20, respectively. Explain your findings.
Finally, discuss how sample size impacts the results and what this implies about the role of randomization for selection bias.