---
title: "Take Home Exercise 1"
subtitle: "Replicating an Experimental Study (THE 1)"
author: "Nicolas Waser"
date: "2024-11-14"
format:
  pdf:
    toc: false
    number-sections: true
    number-depth: 3
    colorlinks: true
editor: 
  markdown: 
    wrap: 72
---

***Exercise 1:** Summary of the Paper and Main Findings Provide a brief
summary of the paper you are replicating. Describe the main findings,
especially those related to Study 1.*

*Disclaimer. For this exercise I made use of the following AI Generation
tools: Deepl.com for translation purposes, the integrated Copilot tool
in R Studio for generating code and text and the free versions of
OpenAI's GPT-4o Mini, Anthropics Claude 3.5 Sonnet and Mistral's
Codestral AI model for generating code and assisting me in solving the
tasks.*

The authors posit that according to the Homonationalist thesis, support
for LGBT+ issues among anti-immigration nativists and right-wing
populists is not entirely genuine (organic liberalism thesis) and is
instead used strategically (instrumental liberalism thesis) as a tool to
justify further exclusion of immigrants, especially Muslims, who are
perceived to be more homophobic, than the average western citizen. The
rational behind this could be summed up as: "The enemy of my enemy is my
friend" or more drastically: "I hate them (Muslims) more than I hate the
other group (LGBT+ people)". In order to test this hypothesis, the
authors conducted two survey experiments in the UK and Spain. The
dependent variable is "Support for LGBT+ inclusive education in school"
and was measured by means of survey response on an order randomized
scale from 0 to 11, but then for the sake of simplicity were turned into
a dichotomous variable for support (0 or 1).

In the UK, for study 1, the experimental research design randomly
assigned 1200 participants into a control group and a treatment group.
Those in the control group were exposed to text and pictures of
nondenominational people with conventional British names protesting
against LGBT+ inclusive education in schools. The treatment group was
exposed to the same pictures and text, but the names and pictures were
changed to appear Muslim. The authors expected that individuals who
harbor negative predispositions toward immigration in the treatment
group, will display more socially liberal preferences toward LGBT+
education than individuals with comparable anti-immigrant sentiment in
the control group. Indeed, the results showed that (with an average
treatment effect of 9.7% points or a 22% increase against the baseline)
respondents with an anti-immigrant disposition in the treatment group
showed increased support for LGBT+ inclusive education in schools than
comparable individuals in the control group. On the other hand, among
those with positively predisposed toward immigration, treatment exhibits
no significant effect below the 5% significance level. But overall, it
can be said, that support for LGBT+ inclusive education is positively
correlated with liberal attitudes towards immigrants.

In Study 2, conducted in Spain, chosen because of generally more
positive attitude levels towards immigration and LGBT+ issues than in
the UK, and a right-wing populist party (VOX), that unlike its
counter-part in the UK (UKIP), is not following a homonationalist
strategy and is more openly outspoken against LGBT+ issues. They found
that regardless of immigration sentiments, treatment resulted in an
10-11% increase in support for LGBT+ inclusive education in schools.
However, the effect was stronger among those with anti-immigrant
sentiments (21% vs. 14%), when compared against the baseline.
Furthermore, the authors also looked at shifts in citizens pride in
"Western Culture". Only among individuals with nativist's views could
they reasonably demonstrate increased levels of "pride in Western
Culture" in the treatment group.

With these findings in mind, the authors argue that the homonationalist
thesis is correct and that the increased levels of support for LGBT+
issues among the wider general public in western countries is not
entirely genuine, but also in part the result of a strategic move by
right-wing populists to deradicalize their image by defining themselves
by "who they are not" rather than "who they are". They further speculate
that as soon as the political context changes and LGBT+ rights are at
stake (but absent of the perceived threat of "immigrants"), the support
for LGBT+ issues among right-wing populists will decrease.

***Exercise 2:** Data Preparation and Exploration Use the data file
study1_data.csv to begin the replication process. Identify and describe
the experimental variables (i.e., treatment and immigration attitudes)
and provide visualizations of their distribution. Then, select up to
four covariates (e.g., gender, age, etc.) and plot their distribution
too. If necessary, clean or transform variables. Document any changes.*

From looking at the csv table for Study1 as well as looking at the
study1.R script, one can deduce which variables are likely of interest
for the replication. As a first step, I need to find the dependent
variable (DV) support for LGBT+ inclusive education, which according to
the text had a 0-11 range with a randomized scale order that was
transformed to 0 and 1 for easier interpretation. Secondly, I need to
find the experimental variables for treatment and immigration attitudes.
The treatment variable concerns the exposure to Muslim names and
pictures and is most likely operationalized as a binary variable (0,1).
The variable for immigration attitudes is likely to be operationalized
as an ordinal scale (0-10) or as a binary variable.

A look at the Csv table revealed the following variables as of interest
for the replication:

-   Treatment variable: outcome_treat (categorical with NAs), treatment
    (ordinal)

-   Immigration attitudes: imm_1, imm_2, (both ordinal) imm1mean
    (continuous, same value), immbelow, imm3 ( both ordinal)

-   Dependent Variable: support (ordinal), support2 (categorical)

By also looking at the study1.R script in the code chunk below, I can
see that the following variables are of interest for the replication,
while the rest can likely be ignored:

```{r}
#df <-df%>% 
  #mutate(treat= as.factor(treatment), 
         #treatnum= as.numeric(treatment), 
         #[...]
         #immbelow= as.factor(immbelow),
         #imm3= as.factor(imm3),
         
##Figure 3## 
#model1<- glm (support ~ treat*imm_1, data=df, family="binomial")
#summ(model1, robust=TRUE)
#df$predictINT1<-predict(model1, df, type="response")

#treat <- subset(df, treatnum==1)
#control <- subset(df, treatnum==0)
#proimm <- subset(df, immbelow==0)
#noproimm <- subset(df, immbelow==1)

##FIGURE 4##
#modelsub1<- lm (support ~ treat, data=proimm)
#summ(modelsub1, robust=TRUE)
#proimm$predictedb<-predict(modelsub1, proimm)

#modelsub2<- lm (support ~ treat, data=noproimm)
#summ(modelsub2, robust=TRUE)
#noproimm$predictedb<-predict(modelsub2, noproimm)

#treatsub1 <- subset(proimm, treatnum==1)
#controlsub1 <- subset(proimm, treatnum==0)
#treatsub2 <- subset(noproimm, treatnum==1)
#controlsub2 <- subset(noproimm, treatnum==0)
```

-   Treatment variable: treatment is transformed twofold -\> treat
    (converted to cat. variable for model transformation), treatnum
    (converted to numeric binary variable for graph)

-   Immigration attitudes: immbelow (cat., converted to binary variable
    for graph), imm_1 (ordinal, 0-10 with 10 standing for maximal
    positive attitude towards immigrants)

-   Dependent Variable: support (ordinary)

From this I conclude that the following variables are of interest for
the replication:

-   Treatment variable operationalization: treatment (0,1)

-   Immigrant attitudes operationalization: imm_1 (ordinal scale 0-10)
    ((or immbelow, for binary distribution Pro-/AntiImmigration))

-   Dependent Variable operationalization: support (0,1)

Below I will provide visualizations of the distribution of these
variables. Let's start with the treatment variable. But first, I need to
load the data sets and the necessary packages. I will then transform the
treatment variable for better visualization and provide a table and a
bar plot of the distribution of the treatment variable.

```{r}
study1 <- 
  read.csv("/Users/nicolaswaser/New-project-GitHub-first/R/Take-Home-Ex.-1/Input Data/study1_data.csv")
library(tidyverse)
library(ggplot2)
#is.factor(study1$treatment)
study1 <-study1%>% 
  mutate(treatment= as.factor(treatment))
#is.factor(study1$treatment)
```

Variable distribution of "treatment" in the form of a table and a bar plot:
```{r}
table(study1$treatment)
```
```{r}
ggplot(study1, aes(x=treatment)) + geom_bar() + 
  labs(title="Distribution of Treatment Variable",
       x="Treatment", y="Frequency")
```
0 = control group, 1 = treatment group
Unsurprisingly, the treatment variable is distributed evenly across the
two groups, as it should be in a randomized experiment.

Next, I will visualize the distribution of the Immigration attitude variable "imm_1"::
```{r}
table(study1$imm_1)
```
```{r}
ggplot(study1, aes(x=imm_1)) + geom_bar() +
  labs(title="Distribution of attitudes towards immigrants",
       x="Attitudes towards immigrants (from negative to positive)",
       y="Frequency")
```
Ordinal scale: 0 = most negative attitude towards immigrants, 1 = most positive attitude towards immigrants 
The distribution of attitudes towards immigrants is skewed towards the positive side, with more respondents having a positive attitude.

Lastly, I will visualize the distribution of the dependent variable "support for LGBT+ inclusive education":
```{r}
table(study1$support2)
```
```{r}
ggplot(study1, aes(x=support2)) + geom_bar() + 
  labs(title="Distribution of support for LGBT+ inclusive education", 
       x="Stance on LGBT+ inclusive education", y="Frequency")
```

The distribution of support for LGBT+ inclusive education is skewed towards the positive side, with most respondents supporting LGBT+ inclusive education.

Selecting covariates - I pick the following covariates for the analysis:
- age: Age is often associated with a change in political attitudes over time -> age (numeric) or agecat (categorical) -> I pick age so that I have a continuous variable in the analysis

- gender: Gender is associated with different attitudes towards LGBT+ rights and immigration and could impact that relationship as well -> I pick the binary variable gender (categorical, Man, Woman)

- sexual identity/orientation: I expect that identifying as LGBT+ will have a big influence on the relationship of the experimental variables and the DV. -> I pick the variable queer (categorical, Heterosexual, Not heterosexual) 

- race: not being white (in the case of the UK) might also impact the relationship between IV's and the DV. I also wanted to add a non-binary categorical variable -> I pick the variable race.

Age:
```{r}
table(study1$agecat)
```
Data Cleaning: There are 3 Nas in the data frame that affect all the covariates I selected. I will remove from the data set:
```{r}
nas_by_column <- 
  colSums(is.na(
    study1[, c("age", "agecat", "gender", "queer", "race")]))
#print(nas_by_column)
rows_with_nas <- 
  which(rowSums(is.na(
    study1[, c("age", "agecat", "gender", "queer")])) > 0)
#print(rows_with_nas)
study1_cleaned <- study1[-rows_with_nas, ]
#print(nrow(study1))
#print(nrow(study1_cleaned))
```
```{r}
ggplot(study1_cleaned, aes(x=age)) + 
  geom_histogram(binwidth = 1) + scale_x_continuous() + 
  labs(title="Distribution of age", x="Age", y="Frequency")
```
Also looking at age cohorts:
```{r}
ggplot(study1_cleaned, aes(x=agecat)) + geom_bar() +
  labs(title="Distribution of age cohorts", x="Age cohort", 
       y="Frequency")
```
The distribution of age is skewed towards older respondents, with most respondents being in their 30s and above.

Gender:
```{r}
table(study1_cleaned$gender)
```
```{r}
ggplot(study1_cleaned, aes(x=gender)) + geom_bar() + 
  labs(title="Distribution of gender", x="Gender", y="Frequency")
```
There are slightly more Men than Women in the sample, but overall they are distributed evenly.

Sexual Identity:
```{r}
table(study1_cleaned$queer)
```
```{r}
ggplot(study1_cleaned, aes(x=queer)) + geom_bar() + 
  labs(title="Distribution of sexual identity", x="Sexual identity",
       y="Frequency")
```
The large majority of respondents identify as heterosexual, with only a small minority identifying as not heterosexual.

Race:
```{r}
table(study1_cleaned$race)
```
```{r}
ggplot(study1_cleaned, aes(x=race)) + geom_bar() + 
  labs(title="Distribution of different races", x="Race", y="Frequency")
```
Overwhelming majority of respondents identify as white, with the rest distributing evenly.

Exercise 3: Covariate Balance
Check for balance across covariates and report the results in a table and a plot.
Explain your findings. Why would you expect randomization to lead to balance across covariates?

```{r}
study1_cleaned_transformed <-study1_cleaned%>% 
  mutate(treatment= as.factor(treatment))
```

```{r}
# Checking for balance across covariates
# Loading necessary libraries
library(dplyr)
library(ggplot2)
install.packages("tableone")
library(tableone)
library(knitr)

# Specifying the treatment variable and covariates
treatment_var <- "treatment"
covariates <- c("gender", "age", "queer", "race")

# Creating a balance table using 'CreateTableOne' from the tableone package
# This will show means and standard deviations (or medians, counts) by treatment group
balance_table <- CreateTableOne(vars = covariates, strata = treatment_var, data = study1_cleaned_transformed, test = TRUE)

print(balance_table, smd = TRUE)  # SMD = standardized mean differences
```

